Secure Composition \\ of Robust and Optimising Compilers

### 2023-09-27 csf/main.tex
Change text
Additions to seqcompo

Any technique to combine existing monitors $T_A$, $T_B$ (that check for properties $A$ and $B$) into one $T_{AB}$ (that checks for both $A$ and $B$) is sufficient for the sake of the analysis of this paper.
Note that monitors only apply to trace-properties, but not, e.g., relational hyperproperties~\cite{abate2019jour}.
\MP{what}

@@

Our main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
% This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
% \MP{what deos this mean?}
The \bul{sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$} is defined as follows:
Given an $\src{L}$ program $\src{p}$ and compilers $\cc{\src{L}}{\trg{L}},\cc{\trg{L}}{\obj{L}}$, \rul{its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$}.
\begin{definition}
  \bul{$\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$} $\isdef$ Given $\src{p}$, yield \rul{$\cc{\trg{L}}{\obj{L}}\left(\cc{\src{L}}{\trg{L}}(\src{p})\right)$}
\end{definition}

@@

Moreover, while this paper assumes a unified trace-model, it is merely a technical exercise to allow for different trace-models between programming languages~\cite{abate2021extacc}.

\MP{moved this}
Note that the theorem may not be applicable in certain cases.
For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
By \Cref{def:trace:tmsdef}, this violates $\tmssafe$.
Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
This, however, is a requirement of the preservation statement (the property of interest must be in $\class[_1]\cap\class[_2]$).

>>>>

Any technique to combine existing monitors $T_A$, $T_B$ (that check for properties $A$ and $B$) into one $T_{AB}$ (that checks for both $A$ and $B$) is sufficient for the sake of the analysis of this paper.

@@

Our main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
% This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
% \MP{what deos this mean?}
The \bul{sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$} is defined as follows:
Given an $\src{L}$ program $\src{p}$ and compilers $\cc{\src{L}}{\trg{L}},\cc{\trg{L}}{\obj{L}}$, \rul{its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$}.
\begin{definition}[Sequential Composition]
  \bul{$\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$} $\isdef$ Given $\src{p}$, yield \rul{$\cc{\trg{L}}{\obj{L}}\left(\cc{\src{L}}{\trg{L}}(\src{p})\right)$}
\end{definition}

@@

So far, this paper assumes a unified trace-model and secure compilers that do not change the behavior.
It is worth noting that in this setting, the theorem may not be applicable.
For example, consider $\rtp{\cc{\src{L}}{\trg{L}}}{\tmssafe}$ and $\rtp{\cc{\trg{L}}{\obj{L}}}{\smssafe}$, where $\cc{\trg{L}}{\obj{L}}$ introduces code that allocates a table that stores bounds information of pointers, but fails to insert code that deallocates this table.
By \Thmref{def:trace:tmsdef}, this violates $\tmssafe$.
Therefore, no behavior resulting from programs produced by the $\cc{\trg{L}}{\obj{L}}$ compiler is secure with respect to $\tmssafe$, but only for $\smssafe$.
$\cc{\trg{L}}{\obj{L}}$ modifies the behavior and because of this, the theorem is inapplicable.

This section re-introduces secure compiler composition without this limitation. 
To this end, it utilises existing work on secure compilation~\cite{abate2021extacc} that defines Galois connections for several real-world compiler scenarios, such as a compilation pass that exploits undefined behavior.
This section is concerned about source-level properties, and henceforth uses the existential image of a source-level property $\src{\pi}$ to map $\src{\pi}$ to a target-level property:

\begin{definition}[Existential Image of Properties]\label{def:eximg:prop}
  Given a cross-language trace relation $\sim$,
  $$
  \tau_\sim(\src{\pi}) \isdef \left\{
    \trg{\trace} \mid \exists\src{\trace}.\src{\trace}\sim\trg{\trace} \text{ and }
    \src{\trace}\in\src{\pi}
  \right\}
  $$
\end{definition}

\Thmref{def:eximg:prop} can be lifted to hyperproperties/classes as follows:
\begin{definition}[Existential Image of Hyperproperties]\label{def:eximg:hprop}
  Given a cross-language trace relation $\sim$,
  $$
  \tilde{\tau}_\sim(\src{\Pi}) \isdef \left\{
    \trg{\pi} \mid \exists\src{\pi}.\src{\pi}\in\src{\Pi} \text{ and }
    \trg{\pi}=\tau_\sim\left(\src{\pi}\right)
  \right\}
  $$
\end{definition}

\bul{A cross-language relation $\sim$ is well-formed with respect to $\src{\Pi}$} iff \rul{the existential images preserve membership}.
\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Concrete Property $\src{\Pi}$]\label{def:wf:tracerel}
  $$
  \text{\bul{$\wf{\sim}{\src{\Pi}}$}} \isdef \text{\rul{$\forall \src{\pi}\in\src{\Pi}, \tau_\sim(\src{\pi})\in\tilde{\tau}_\sim(\src{\Pi})$}}
  $$
\end{definition}
\begin{definition}[Well-formedness of Cross-Language Trace Relations for a Class $\src{\class}$]\label{def:wfc:tracerel}
  $$
  \wfc{\sim}{\src{\class}} \isdef \forall \src{\Pi}\subseteq\src{\class}, \tau_\sim(\src{\Pi})\subseteq\tilde{\tau}_\sim(\src{\class})
  $$
\end{definition}
\begin{lemma}[Well-Formedness Equivalence for Classes]\label{lem:wf:equiv:wfc}
  $\;\;$
  $ \wf{\sim}{\src{\class}} \Leftrightarrow\ \wfc{\sim}{\src{\class}} $\Coqed
\end{lemma}

\Thmref{def:rtp} needs to be adapted to account for different target-level trace-models:

\begin{definition}[Robust Compilation with $\sim$]\label{def:rtpsim}
  $$\rtpsim{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim} \isdef
  \forall \src{\pi} \src{p}.
    \text{if }\rsat{\src{\progvar}}{\src{\pi}},
    \text{then }\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tilde{\tau}_\sim\left(\src{\pi}\right)}.
  $$
\end{definition}

To state a similar theorem as \Thmref{thm:rtp} but for \Cref{def:rtpsim}, it is necessary to define composition of cross-language trace relations:
\begin{definition}[Composing Cross-Language Trace Relations]
  Given $\sim_1 : \src{\trace}\to\irl{\trace}\to\text{Prop},\sim_2 : \irl{\trace}\to\trg{\trace}\to\text{Prop}$, then 
  $$  
  \sim_1\bullet\sim_2\ \isdef \lambda \src{\trace}\ \trg{\trace}. \exists \irl{\trace}. \src{\trace}\sim_1\irl{\trace} \text{ and } \irl{\trace}\sim_2\trg{\trace}
  $$
\end{definition}

\begin{theorem}[Sequential Composition of Secure Compilers]\label{thm:rtpsim}
  If $\rtpsim{\cc{\src{L}}{\trg{L}}}{\class[_{1}]}{\sim_1}$, $\rtpsim{\cc{\trg{L}}{\obj{L}}}{\tilde{\tau}_{\sim_1}\left(\class[_{2}]\right)}{\sim_2}$, and $\wf{\sim_1}{\class[_2]}$, then $\rtpsim{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\class[_{1}]\cap\class[_{2}]}{\sim_1\bullet\sim_2}$. \Coqed
\end{theorem}

Reconsidering the example from the beginning of the section, suppose $\sim_{{\smssafe}}$ is a suitable trace relation such that $\trg{\trace}\sim_{{\smssafe}}\irl{\trace}$ establishes the fact that $\irl{\trace}$ contains an allocation of a table containing bounds information which is never deallocated.
It is impossible that $\wf{\sim_{{\smssafe}}}{{\tmssafe}}$, since every behavior of compiled $\trg{L}$ programs never contains a deallocation of the bounds-information table.
Still, supposing $\rtpsim{\cc{\src{L}}{\trg{L}}}{\tmssafe}{{\sim_{\tmssafe}}}$, $\rtpsim{\cc{\trg{L}}{\irl{L}}}{\smssafe}{{\sim_{\smssafe}}}$,
and $\wf{{\sim_{\tmssafe}}}{{\smssafe}}$ allows to apply \Thmref{thm:rtpsim}.
However, the resulting class $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$, which the following lemma shows to hold in general.

\begin{lemma}[Composition is Useless]\label{lem:useless}
  If $\wf{\sim_1}{\class}$, $\neg\left(\wf{\sim_2}{\tilde{\tau}_{\sim_1}\class}\right)$, then $\tilde{\tau}_{\sim_1\bullet\sim_2}\left(\class\right)=\emptyset$. \Coqed
\end{lemma}

With this, the previous claim follows as a simple corollary:
\begin{corollary}
  $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\cap\smssafe\right)=\emptyset$.
\end{corollary}
\begin{proof}
  Note that $\neg\left(\wf{\sim_{{\smssafe}}}{{\tmssafe}}\right)$, since the bounds-table is never deallocated and therefore, for any $\trg{\trace}\in\tmssafe$, $\tau_{\sim_{\tmssafe}}\left(\trg{\trace}\right)\notin\tilde{\tau}_{\sim_{\tmssafe}}\left({\tmssafe}\right)$.
  However, by assumption, $\wf{\sim_{{\tmssafe}}}{{\tmssafe}}$.
  Using \Thmref{lem:useless} it follows that $\tilde{\tau}_{{{\sim_{\tmssafe}}\bullet{\sim_{\smssafe}}}}\left(\tmssafe\right)=\emptyset$.
\end{proof}
